Why Did Russia Fare So Poorly during COVID: An Analysis of Why One of the Best Prepared Countries Could Not Control Excess Deaths

### 2024-02-13 paper/paper.qmd
Temp Push
Temp Push

---
title: "Why Did Russia Fare So Poorly during COVID: An Analysis of Why One of the Best Prepared Countries Could Not Control Excess Deaths"
author: 
  - Aryaman Sharma
  - Aviral Bhardwaj
  - Janel Gilani
thanks: "Code and data are available at: https://github.com/Ary4m3n/covid-effect-russia.git; Replication on Social Science Reproduction platform available at: "
date: today
date-format: long
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean tempus libero sit amet nisl dapibus, varius dictum magna pellentesque. In lacinia nec ligula eget condimentum. Aenean imperdiet enim sit amet felis volutpat, eu ornare ante pretium. Donec ullamcorper tempus facilisis. Mauris porttitor libero ultrices tortor facilisis, quis varius sapien ultricies. Pellentesque et metus vel quam mollis dignissim ultricies a dui. Cras sodales porttitor pulvinar. Proin vel nunc sit amet sapien venenatis pellentesque nec at felis."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(janitor)
library(dplyr)
library(rnaturalearth)
library(sf)
library(knitr)
```

@@

This paper will follow a reproduction of Jennifer B. Nuzzo and Jorge R. Ledesma's paper [@citeRepPaper] and findings and apply a Russian-facing lens to discuss the impact of the COVID-19 outbreak on Russia and what could have been done better and can be improved for the future in order for the impact of any upcoming pandemics can be lessened. Our paper aims to replicate their claim of how well-prepared countries did so poorly during the COVID-19 outbreak, where the Nuzzo and Ledesma have concentrated solely on the United States but in this paper we look at Russia with the same regards and analyse the impact of the pandemic on the death toll, and finally provide for ways to improve the laws, regulations and plans to tackle such outbreaks in the future. Our reproduction found out that how ever well prepared Russia seemed to be, it experienced an unprecedented rise in the death rate compared to other countries, and its excess death rate was one of the highest amongst all countries. This leads us to make an important judgement that better planning and regulations can help reduce the impact of such outbreaks in the future, which will be studied in this paper.

This paper is structured using the following sections: Data, Results and Discussion. In the Data (@sec-data) section, the data source of the datasets from the paper being replicated [@citeRepPaper] is discussed and the measurement and data cleaning process is outlined. In the Results (@sec-results) section, the paper summarizes the data findings and presents relevant maps and plots in order to study the trends. The paper ends with the Discussion (@sec-discussion) section, where the findings of the paper have been analysed and delved deeper into by studying the measures that can be taken to prevent the country from experiencing such an extensive effect of future pandemics. Additionally, the limitations and a further scope for the paper has been discussed here.

@@

As mentioned above, the data analysed in this paper is from the reproduction package of the paper by Nuzzo and Ledesma [@citeRepPaper]. The data was cleaned and analysed using the open source `R` programming language [@citeR]. R libraries and packages such as `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `ggplot2` [@citeGgplot], `knitr` [@citeKnitr], `readr` [@citeReadr], `dplyr` [@citeDplyr], `rnaturalearth` [@citeRnaturalearth] and `sf` [@citeSf]. In the following sections, we will discuss the raw data in (@sec-data-raw), the data cleaning process in (@sec-data-clean) and then move on to discussing the measurement in (@sec-data-measurement).

@@

The first raw dataset includes data on the Global Health Security index for all countries. The GHS index is an indicator of how well prepared or "secure" a country is against epidemics or pandemics like the COVID-19 outbreak. This dataset has a plethora of data on indices that are not relevant to us for this paper. Additionally, the dataset contains data for years 2019 and 2021. For our analysis, we only require data for the GHS index of countries for 2021.

The second raw dataset includes data on all cause death rates for all countries from 1960-2021. This dataset has loads of information that was mainly unnecessary to us for this paper, mainly the data from 1960-2010 which is not required as we are only interested in looking at the effect of COVID-19 which is mainly data from 2020-2021. However, in order to study the trend, we will use data for all cause death rates for countries around the world from 2010-2021.

The third raw dataset is from the Institute of Health Metrics and Evaluation (IHME) which presents data on the COVID death rate, excess deaths, excess death rate and ratio of excess death rate over COVID death rate for all countries in the world. Again, this dataset too contained unnecessary information that was of no need to us which the cleaned data process outlined in @sec-data-clean.

@@

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean tempus libero sit amet nisl dapibus, varius dictum magna pellentesque. In lacinia nec ligula eget condimentum. Aenean imperdiet enim sit amet felis volutpat, eu ornare ante pretium. Donec ullamcorper tempus facilisis. Mauris porttitor libero ultrices tortor facilisis, quis varius sapien ultricies. Pellentesque et metus vel quam mollis dignissim ultricies a dui. Cras sodales porttitor pulvinar. Proin vel nunc sit amet sapien venenatis pellentesque nec at felis.

>>>>

---
title: "Why Did Russia Fare So Poorly during COVID: An Analysis of Why One of the Best Prepared Countries Could Not Control Excess Deaths"
author: 
  - Aryaman Sharma
  - Aviral Bhardwaj
  - Janel Gilani
thanks: "Code and data are available at: https://github.com/Ary4m3n/covid-effect-russia.git; Replication on Social Science Reproduction platform available at: "
date: today
date-format: long
abstract: "GHS Index"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
# install.packages("rnaturalearthdata")
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(janitor)
library(dplyr)
library(rnaturalearth)
library(sf)
library(knitr)
```

@@

This paper will follow a reproduction of Jennifer B. Nuzzo and Jorge R. Ledesma's paper [@citeRepPaper], where they used The Global Health Security (GHS) Index an indicator of how well prepared or "secure" a country is against epidemics or pandemics like the COVID-19 outbreak. We in-turn will use these findings and apply a Russian-facing lens to discuss the impact of the COVID-19 outbreak on Russia and what could have been done better and can be improved for the future in order for the impact of any upcoming pandemics can be lessened. Our paper aims to replicate their claim of how well-prepared countries did so poorly during the COVID-19 outbreak, where the Nuzzo and Ledesma have concentrated solely on the United States but in this paper we look at Russia with the same regards and analyse the impact of the pandemic on the death toll, and finally provide for ways to improve the laws, regulations and plans to tackle such outbreaks in the future. Our reproduction found out that how ever well prepared Russia seemed to be, it experienced an unprecedented rise in the death rate compared to other countries, and its excess death rate was one of the highest amongst all countries. This leads us to make an important judgement that better planning and regulations can help reduce the impact of such outbreaks in the future, which will be studied in this paper.

This paper is structured using the following sections: Data, Results and Discussion. In the Data (@sec-data) section, the data source of the datasets from the paper being replicated [@citeRepPaper] is discussed and the measurement and data cleaning process is outlined. In the Results (@sec-results) section, the paper summarizes the data findings and presents relevant maps and plots in order to study the trends. The paper ends with the Discussion (@sec-discussion) section, where the findings of the paper have been analysed and delved deeper into by studying the measures that can be taken to prevent the country from experiencing such an extensive effect of future pandemics. Additionally, the limitations and a further scope for the paper has been discussed here and concluded in the end.

@@

As mentioned above, the data analysed in this paper is from the reproduction package of the paper by Nuzzo and Ledesma [@citeRepPaper]. The data was cleaned and analysed using the open source `R` programming language [@citeR]. R libraries and packages such as `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `ggplot2` [@citeGgplot], `knitr` [@citeKnitr], `readr` [@citeReadr], `dplyr` [@citeDplyr], `rnaturalearth` [@citeRnaturalearth] and `sf` [@citeSf]. Since, we have replicated the process of data cleaning and analysis, the data used in this paper is the same as the data used in the original paper.

In the following sections, we will discuss the raw data in (@sec-data-raw), the data cleaning process in (@sec-data-clean) and then move on to discussing the measurement in (@sec-data-measurement).

@@

The first raw dataset includes data on the Global Health Security index for all countries. This dataset has a plethora of data on indices that are not relevant to us for this paper. Additionally, the dataset contains data for years 2019 and 2021. For our analysis, we only require data for the GHS index of countries for 2021.

The second raw dataset includes data on all cause death rates for all countries from 1960-2021. Within this dataset, a plethora of information exists, the majority of which proved extraneous for the purposes of our paper. Specifically, data preceding the year 2010 is deemed unnecessary for our analysis, since we are only focusing on COVID-19 we will use this dataset to study the trend in death rates for countries around the world from 2010-2021.

The third raw dataset originates from the Institute of Health Metrics and Evaluation (IHME) and provides information on COVID death rates, excess deaths, excess death rates, and the ratio of excess death rates to COVID death rates for all countries worldwide. Similar to the previous datasets, this dataset also contains extraneous information that is irrelevant to our study. However, we addressed this issue through the data cleaning process outlined in Section @sec-data-clean.

@@

When discussing COVID-19, (1) issues with underreported or missing data, particularly related to non-responses, are prevalent, (2) the lack of a standardized definition of COVID-19 deaths, which can lead to measurement errors (3) vastness of countries poses logistical challenges in reporting and conducting surveys, making it difficult to gather comprehensive data (4) variations in testing methodologies and availability of resources across regions can further distort the data.

Many countries, especially developing nations like India, Brazil, and Russia, poses above issue. This poses a significant challenge in accurately measuring the pandemic's impact on a country. For example, the excess death rate, a variable in our analysis, measures the number of deaths occurring in a specific period above the expected number. This metric helps estimate the pandemic's impact. However, due to data truncation, the actual death rate isn't always precise, leading to inaccuracies in calculating the excess death rate. This introduces measurement errors, undermining the accuracy of both the actual death rate and the excess death rate.

With respect to measurement and accounting for the GHS index, it encompasses various measurements grouped into six broad categories: (1) prevention, (2) detection and reporting, (3) rapid response, (4) health system, (5) compliance with international norms, and (6) risk environment. However, various measurement errors and qualitative biases can arise from the subjective nature of these categories. For example, the "prevention" category includes measurements such as "vaccination coverage" and "biosecurity and biosafety." However, the definition of "biosecurity and biosafety" can vary across countries, leading to measurement errors. Additionally, the GHS index is based on self-reported data, which can be subject to biases and inaccuracies. This can lead to measurement errors in the GHS index, which can further distort the analysis.
